---
title: "Node order matters: comparative analysis of soil water limitation effects on root anatomy between nodal position and maize genotypes (Zea mays L.)"
output: github_document
---

Tina Koehler, Yunhee Kim, Shu-Yin Tung, Adrien Heymans, Nicolas Tyborski, Franziska Steiner, Andreas J. Wild, Johanna Pausch, Mutez A. Ahmed, Hannah Schneider

This Rmd is part of the RootDiversity repository and branched for the above mentioned article. 

## GRANAR-MECHA: modeling pipeline to estimate emergent hydraulic properties"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### loading dependencies
library(tidyverse)
library(plyr)
library(deldir)
library(alphahull)
library(xml2)
library(sp)
library(viridis)
library(readxl)
`%!in%` <- compose(`!`, `%in%`)
setwd("~/GitHub/Tina2024/")

```

```{r granar, include=FALSE}
#### loading GRANAR 
source("./GRANAR/R/granar.R")
source("./GRANAR/R/micro_hydro.R")
### loading mock parameter file
params <- read_param_xml("./GRANAR/www/Zea_mays_CT.xml")
```

### Used known model: 
- The Generator of root anatomy in R - **GRANAR** [Heymans et al. 2020](https://doi.org/10.1104/pp.19.00617)
- The model of explicit hydraulic anatomy - **MECHA** [Couvreur et al. 2018](https://doi.org/10.1104/pp.18.01006)

## Run GRANAR to simulate cross-section anatomy from trait values

The Generator of root anatomy (GRANAR; Heymans et al. 2020) use a set of input that can be derive from a few anatomical traits. 
The anatomical traits needed to simulate root cross-section anatomy are:

| Anatomical parameter          | Unit |
|-------------------------------|------|
| Root radius                   | mm   |
| Stele radius                  | mm   |
| Cortex file number            | -    |
| Aerenchyma area               | mm^2 |
| Metaxylem area                | mm^2 |
| Nbr of metaxylem elm          | -    |

```{r generate_anatomies, eval=FALSE}

#### loading input from anatomical trait dataset
Raw_data <- readxl::read_excel("./www/Root no CR6_update_2.xlsx")

Raw_data[Raw_data$Treatment=='Drought',2]<-'sheltered'
Raw_data[Raw_data$Treatment=='Well watered',2]<-'non-sheltered'



Sampl = Raw_data%>%
  mutate(RXA = `Root area`,
         TSA = `Stele area`,
         TCA = `Cortex area`,
         AA = `Aerenchyma area`,
         aerenchyma= `Aerenchyma percent`/100,
         radius = sqrt(RXA/pi),
         r_stele = sqrt(TSA/pi),
         MXA = `Total Metaxylem area`,
         nX = `Metaxylem number`,
         X_size = 2*sqrt((MXA/nX)/pi),
         CF = `Cortical file number`,
         OneC = (radius-r_stele)/(CF+2),
         OC = 2*sqrt(`Cortical cell size`/pi),
         ratio = (2+0.07456*r_stele*1000)/nX,
         nPX = nX*ratio,
         PXA_1 = 1000^2*(sqrt(radius/35)/10)^2,
         k_protxyl_s = PXA_1^2/(8*pi*200*1E-5/3600/24)*1E-12,
         kx_unM = k_protxyl_s*nPX*200/1E4, # kx when only the proto xylem have their cell wall lignified 
         LMXA = MXA,
         LMXA_1 = LMXA*1000^2/nX,
         k_Mxyl_s = LMXA_1^2/(8*pi*200*1E-5/3600/24)*1E-12,
         kx_M = k_Mxyl_s*nX*200/1E4 + kx_unM)

Sampl%>%
  ggplot()+geom_histogram(aes(RXA-TCA), fill = 'blue', alpha = 0.5, bins = 50)+geom_histogram(aes(TSA), fill = 'red', alpha = 0.5, bins = 50)+
  theme_classic()+xlab('TCA [mm2]')

av_dat = Sampl%>%
  dplyr::group_by(Genotype, Treatment, Roottype)%>%
  dplyr::summarise(m_CF = mean(CF, na.rm = T),
                   m_OC = mean(OC, na.rm = T),
                   m_aerenchyma = mean(aerenchyma, na.rm = T), .groups = "drop")

Sampl = left_join(Sampl, av_dat , by = c("Genotype", "Treatment", "Roottype"))

Sampl = Sampl%>%
  mutate(CF = ifelse((is.na(CF) | CF <= 4), m_CF,CF),
         OneC = (radius-r_stele)/CF,
         OC = ifelse((is.na(OC) | OC > 0.2), m_OC,OC),
         aerenchyma = ifelse(is.na(aerenchyma), m_aerenchyma,aerenchyma))%>%
  arrange(radius)

Sampl$id = 1:nrow(Sampl)
Sampl %>%
  ggplot(aes(y = id, yend = id))+
  geom_segment(aes(x= 0, xend = r_stele), colour = 'red')+
  geom_segment(aes(x = r_stele, xend = radius ), colour = "blue")+theme_classic()+
  geom_point(aes(x = r_stele + OneC))+
  xlab('cross section radius [mm]')+
  ylab('cross section id')

Sampl$model_RXA = NA
Sampl$model_TSA = NA
Sampl$model_XVA = NA
Sampl$model_AA = NA

###### Proc for the generation of the anatomies
for(i in 1:nrow(Sampl)){
  # Set the input line
  tmp_sampl <- Sampl[i,]
  if(is.na(tmp_sampl$aerenchyma)){
    tmp_sampl$aerenchyma = 0
  }
  # clear results
  if(file.exists("./MECHA/cellsetdata/current_root.xml")){
    file.remove("./MECHA/cellsetdata/current_root.xml")
    file.remove("./MECHA/Projects/GRANAR/in/Maize_Geometry_aer.xml")
  }
  # Run GRANAR and change the parameter for the selected simulation
  sim <- run_granar(params, tmp_sampl)
  # Write the outputs
  file.copy("./MECHA/cellsetdata/current_root.xml", 
            paste0("./MECHA/cellsetdata/root_",i,".xml"), overwrite = T)
  file.copy("./MECHA/Projects/GRANAR/in/Maize_Geometry_aer.xml",
            paste0( "./MECHA/Projects/GRANAR/in/Maize_Geometry_aer_",i,".xml"), overwrite = T)
  
}


## Check for validity


fls <- list.files("./MECHA/cellsetdata/")
fls <- fls[grepl("root_", fls)]

for(j in fls){
  out = granar_metadata(paste0("./MECHA/cellsetdata/",j))
  i <- parse_number(j)
  Sampl$model_RXA[i] = out$value[out$type == "layer_area" & out$param == "all"]
  Sampl$model_TSA[i] = out$value[out$type == "layer_area" & out$param == "stelar"]
  Sampl$model_XVA[i] = out$value[out$type == "layer_area" & out$param == "metaxylem"]
  mAA = out$value[out$type == "layer_area" & out$param == "aerenchyma"]
  Sampl$model_AA[i] = ifelse(length(mAA)>0, mAA, 0)
}


area = rbind(tibble(M_area = Sampl$RXA, S_area = Sampl$model_RXA, what = "Root cross-section"),
             tibble(M_area = Sampl$TSA, S_area = Sampl$model_TSA, what = "Stele"),
             tibble(M_area = Sampl$MXA, S_area = Sampl$model_XVA, what = "Meta Xylem"),
             tibble(M_area = Sampl$AA, S_area = Sampl$model_AA, what = "Aerenchyma"))

area%>%
  ggplot()+
  geom_point(aes(M_area, S_area, colour = what))+
  geom_abline(slope = 1, linetype = 2)+
  theme_classic()+
  facet_wrap(~what, scales = "free")+
  labs(y = "Simulated area [mm²]", x = "Measured area [mm²]")
```


## Run MECHA to estimate emergent hydraulic properties

The model of explicit cross-section hydraulic architecture (MECHA; Couvreur et al. 2018) estimates kr from the root transverse anatomy generated with GRANAR or CellSeT (Pound et al., 2012) and from the subcellular scale hydraulic properties of walls, membranes, and plasmodesmata.

| Sub-cell hydraulic properties   | Unit                 | Value    | Ref.                 |
|---------------------------------|----------------------|----------|----------------------|
| Cell wall conductivity          | $cm^2/hPa/d$         | 0.00024  | Zhu and Steudle 1991 |
| Cell membrane base permeability | $cm/hPa/d$           | 3e-5     | Ehlert et al. 2009   |
| Aquaporin contribution          | $cm/hPa/d$           | 0.00043  | Ehlert et al. 2009   |
| Conductance of plasmodesmata    | $cm^3/hPa/d/plasmo.$ | 5.3e-12  | Couvreur et al. 2018 |
| Cell wall thickness             | $µm$                 | 1.5      | Heymans et al. 2020  |

Three hydraulic scenario were implemeted for each simulation:

- 1. Endodermal Casparian strip 
- 2. Endodermal suberization
- 3. Endodermis full suberization and exodermal Casparian strip.

```{r mecha_root, echo=FALSE}
# ### Proc: estimation of the radial hydraulic conductivities
fls <- list.files("./MECHA/cellsetdata/")
fls <- fls[grepl("root_", fls)]

fl_done = list.files("./MECHA/Projects/GRANAR/out/M1v4/Root/")
fl_done = fl_done[grepl("Macro_prop_1,0_", fl_done)]
ids = parse_number(unlist(str_split(fl_done,"1,0")))
ids <- ids[!is.na(ids)]
fl_done = paste0("root_", ids,".xml")
fls = fls[fls %!in% fl_done]


for(j in fls){
  message("--------------")
  print(j)
  message("--------------")
  
  if(file.exists("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_1,0.txt")){
    file.remove("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_1,0.txt")
    file.remove("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_2,1.txt")
    file.remove("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_4,2.txt")
    file.remove("./MECHA/cellsetdata/current_root.xml")
    file.remove("./MECHA/Projects/GRANAR/in/Maize_Geometry_aer.xml")
  }
  
  # Loading input files for the current estimation
  fc <- file.copy(paste0("./MECHA/cellsetdata/",j), "./MECHA/cellsetdata/current_root.xml", overwrite = T)
  if(fc == FALSE){next()}
  fc <- file.copy(paste0("./MECHA/Projects/GRANAR/in/Maize_Geometry_aer_", parse_number(j), ".xml"),
                  paste0("./MECHA/Projects/GRANAR/in/Maize_Geometry_aer.xml"), overwrite = T)
  if(fc == FALSE){next()}
  
  # MECHA input change
  id <- parse_number(j)
  microhydro(path = "MECHA/Projects/GRANAR/in/Maize_hydraulics.xml",
             kw = 0.00024,
             km = 3e-5,
             kAQP = 0.00043,
             kpl = 5.3e-12)
  
  wallthick(path = "MECHA/Projects/GRANAR/in/Maize_Geometry_aer.xml", 1.5)
  
  # Run MECHA - - - - - - -
  system("python3 ./MECHA/MECHAv4_septa.py")
  message("python script has ended")
  
  # if works well, then:
  if(file.exists("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_1,0.txt")){
    # Save output
    message ("success")
    file.copy("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_1,0.txt",
              paste0("./MECHA/Projects/GRANAR/out/M1v4/Root/Macro_prop_1,0_",id,".txt"), overwrite = T)
    file.copy("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_2,1.txt",
              paste0("./MECHA/Projects/GRANAR/out/M1v4/Root/Macro_prop_2,1_",id,".txt"), overwrite = T)
    file.copy("./MECHA/Projects/GRANAR/out/M1v4/Root/Project_Test/results/Macro_prop_4,2.txt",
              paste0("./MECHA/Projects/GRANAR/out/M1v4/Root/Macro_prop_4,2_",id,".txt"), overwrite = T)
  }else{message ("fail and move to next simulation")}
  
}

```



```{r}

# Read Mecha output
fls <- list.files("./MECHA/Projects/GRANAR/out/M1v4/Root/")
fls <- fls[grepl(".txt", fls)]

K <- tibble(kr = NULL, kx = NULL, sampl_id = NULL, apo = NULL)
for (k in fls){
  M <- read_file(paste0("./MECHA/Projects/GRANAR/out/M1v4/Root/",k))
  tmp_M <- strsplit(M, split="\n")[[1]]
  K_xyl_spec <- as.numeric(strsplit(tmp_M[15], " ")[[1]][5])
  kr_M <- as.numeric(strsplit(tmp_M[17], " ")[[1]][4])
  scenario <- round(parse_number(unlist(str_split(k,"_"))[3])/10)
  sampl_id <- parse_number(unlist(str_split(k,"_"))[4])
  K <- rbind(K, tibble(kr = kr_M, kx = K_xyl_spec, sampl_id = sampl_id, apo = scenario))
}
K = K %>%arrange(sampl_id, apo)

#### Merge the results on the input dataframe
Sampl$Kx = NA
Sampl$kr1_new = NA
Sampl$Kr1 = NA
Sampl$Kr2 = NA
Sampl$Kr3 = NA


# Axial hydraulic conductance (cm3 hPa-1 d-1)
Sampl$Kx[Sampl$id %in% K$sampl_id] = K$kx[K$apo == 1]
Sampl%>%
  filter(!is.na(Kx)) %>% 
  ggplot()+
  geom_point(aes(MXA, Kx), colour = "blue")+
  theme_classic()



# Radial hydraulic conductivity (cm hPa-1 d-1)
Sampl$kr1_new[Sampl$id %in% K$sampl_id] = K$kr[K$apo == 1]
Sampl$kr2_new[Sampl$id %in% K$sampl_id] = K$kr[K$apo == 2]
Sampl$kr3_new[Sampl$id %in% K$sampl_id] = K$kr[K$apo == 4]

# Radial hydraulic conductance (cm3 hPa-1 d-1)
# Multiply by the surface of the segment (1 cm length * perimeter of the root)
Sampl$Kr1[Sampl$id %in% K$sampl_id] = K$kr[K$apo == 1]*2*pi*Sampl$radius[Sampl$id %in% K$sampl_id]/10
Sampl$Kr2[Sampl$id %in% K$sampl_id] = K$kr[K$apo == 2]*2*pi*Sampl$radius[Sampl$id %in% K$sampl_id]/10
Sampl$Kr3[Sampl$id %in% K$sampl_id] = K$kr[K$apo == 4]*2*pi*Sampl$radius[Sampl$id %in% K$sampl_id]/10

write.csv(Sampl, "./Sampl_KrKx.csv")

```

## Radial hydraulic conductance and conductivity

The emerging hydraulic properties are shown in the figures below to highlight the radius effect and the one of the cortex width on the radial hydraulic conductivity and conductance.

```{r Kr_plot, echo = FALSE, warning=FALSE}

Sampl%>%
  filter(!is.na(Kx)) %>% 
  ggplot()+
  geom_point(aes(radius-r_stele, Kr1, colour = radius))+
  theme_classic()+ylab("Radial hydraulic conductivity [cm hPa-1 d-1]")+
  xlab("cortex width [mm]")+
  viridis::scale_colour_viridis()

Sampl%>%
  filter(!is.na(Kx)) %>% 
  ggplot()+
  geom_point(aes(radius-r_stele, kr1_new, colour = radius))+
  theme_classic()+ylab("Radial hydraulic conductivity [cm hPa-1 d-1]")+
  xlab("cortex width [mm]")+
  viridis::scale_colour_viridis()

fit <- nls(kr1_new ~ 1/(a * cortex)^b, data = Sampl%>%filter(!is.na(Kx)) %>% mutate(cortex = radius-r_stele), start = list(a = 1, b = 1))

summary(fit)

fitted_values <- fitted(fit)
residuals <- residuals(fit)
# Calculate RMSE
rmse <- sqrt(mean(residuals^2))

data = Sampl%>%filter(!is.na(Kx)) %>% mutate(cortex = radius-r_stele)

# Calculate R-squared
ss_res <- sum(residuals^2)
ss_tot <- sum((data$kr1_new - mean(data$kr1_new))^2)
r_squared <- 1 - (ss_res / ss_tot)

Sampl%>%
  filter(!is.na(Kx)) %>% 
  ggplot()+
  geom_point(aes(radius-r_stele, kr1_new, colour = radius))+
  geom_line(aes(radius-r_stele, fitted_values), colour = "blue", linetype = 2, size = 1)+
  theme_classic()+ylab("Radial hydraulic conductivity [cm hPa-1 d-1]")+
  xlab("cortex width [mm]")+
  viridis::scale_colour_viridis()


Sampl%>%
  mutate(fcortex = round(radius)) %>% 
  filter(!is.na(Kx)) %>% 
  ggplot(aes(radius-r_stele, Kr1, colour = radius))+
  geom_point(aes(colour = Treatment))+
  geom_smooth(aes(colour= Treatment), method = "lm")+
  theme_classic()+ylab("Radial hydraulic conductivity [cm hPa-1 d-1]")+
  xlab("cortex width [mm]")+
  viridis::scale_colour_viridis(discrete = T)

Sampl%>%
  mutate(fcortex = round(radius)) %>% 
  filter(!is.na(Kx), radius < 5.5) %>% 
  ggplot(aes(radius, (Kr1*10^-6/86400)/0.0001, colour = Treatment))+
  geom_point(aes(colour = Treatment, shape = factor(Roottype)), size = 3, alpha = 0.3)+
  geom_smooth(aes(colour= Treatment), method = "lm", alpha = 0.2)+
  theme_classic()+ylab("Radial hydraulic conductance [m3 MPa-1 s-1]")+
  xlab("Radius [mm]")+
  labs(shape = "Root type")

Sampl%>%
  mutate(fcortex = round(radius)) %>% 
  filter(!is.na(Kx), radius < 5.5) %>% 
  ggplot(aes(Roottype, (Kr1*10^-6/86400)/0.0001, colour = Treatment))+
  #geom_point(aes(colour = Treatment, shape = factor(Roottype)), size = 3, alpha = 0.3)+
  geom_boxplot(aes(colour= Treatment), method = "lm", alpha = 0.2)+
  theme_classic()+ylab("Radial hydraulic conductance [m3 MPa-1 s-1]")+
  xlab("Root type")


Sampl%>%
  mutate(fcortex = round(radius)) %>% 
  filter(!is.na(Kx)) %>% 
  ggplot(aes(Roottype, (Kr1*10^-6/86400)/0.0001, colour = Treatment))+
  #geom_point(aes(colour = Treatment, shape = factor(Roottype)), size = 3, alpha = 0.3)+
  geom_boxplot(aes(colour= Treatment), alpha = 0.2)+
  facet_wrap(~Type)+
  theme_classic()+ylab("Radial hydraulic conductance [m3 MPa-1 s-1]")+
  xlab("Root type")

Sampl%>%
  mutate(fcortex = round(radius)) %>% 
  filter(!is.na(Kx)) %>% 
  ggplot(aes(Roottype, (Kr1*10^-6/86400)/0.0001, colour = Treatment))+
  #geom_point(aes(colour = Treatment, shape = factor(Roottype)), size = 3, alpha = 0.3)+
  geom_boxplot(aes(colour= Treatment), alpha = 0.2)+
  facet_wrap(~Genotype)+
  theme_classic()+ylab("Radial hydraulic conductance [m3 MPa-1 s-1]")+
  xlab("Root type")

```


